A comparative analysis of penalized linear mixed models in structured genetic data

### 2020-09-29 main.tex
Updating paper figs (dichotomous-discordant change) and descriptions

 To evaluate the performance of penalized LMM methods in the presence of these different forms of confounding complexity, we consider environmental effects which are concordant or discordant in nature in the presence of both exponential and dichotomous effect sizes, for a total of four distinct $\bgamma$ patterns. Figure \ref{fig:gamma_structures} provides a visual summary of the four coefficient patterns, where the subpopulations on the x-axis are ordered by genetic similarity so that subpopulation 1 is most genetically similar to subpopulation 2, and least similar to subpopulation 10. This may also be thought of as subpopulations ordered spatially along a 1-dimensional geographic region.
\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.9]{figures/gamma_structure.png}
    \caption{An illustration of four distinct patterns of environmental effects and corresponding $\bgamma$ values in the presence of 10 subpopulations ordered by genetic similarity.}
    \label{fig:gamma_structures}
\end{figure}
\noindent In the case of concordant and dichotomous effects, Figure \ref{fig:gamma_structures} shows that the first five ordered subpopulations are subject to an effect size of 1, while the second five are subject to an effect size of $-$1. In the case of discordant dichotomous effects, the first five subpopulations experience an effect size with an absolute value of \anna{1}, and the second five an absolute effect size of \anna{0.5}. However, the directions of these absolute effects sizes alternate among the ordered subpopulations so that genetically similar subpopulations do not experience similar environmental effects. \anna{not quite 1 and 0.5 because of centering and standardization...how to word this so it's not confusing? - trying better version of dichotomous-discordant...}

@@

We begin by comparing the ordinary Lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied. Estimation accuracy for Lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) and broken down into bias and variance components. Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$. The top row of Figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), Lasso and LMM-Lasso perform comparably. PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that 10 additional predictors (PCs) have been added to the model, but none of them actually explain any variability in the outcome.

The bottom row of Figure \ref{fig:mse} shows the MSE when $\xi = 0.8$. In this case, both PC-Lasso and LMM-Lasso clearly outperform the ordinary Lasso. This improvement is more substantial in the Empirical and Independent Subpopulations data than in the 1D Linear Admixture data.

Qualitatively, this is due to the fact that the Independent Subpopulations and Empirical data sets are characterized by more discrete, nonoverlapping genetic subpopulations, which more directly correspond to the nonoverlapping environmental effects than do the continuous and blended subpopulations of the 1D Linear Admixture data. In other words, a greater proportion of the environmental effect is in the column space of $\boldsymbol{P}_{\bX}$ for the Empirical and Independent Subpopulations data than for the 1D Linear Admixture data. We can quantify the proportion of the environmental component in the column space of $\boldsymbol{P}_{\bX}$ for each data type using the CP metric defined in \eqref{eqn:ratio} of Section \ref{sec:sim_env_conf}. CP $= 0.46, 0.99, 0.64$ for the 1D Linear Admixture, Independent Subpopulations, and Empirical data, respectively. The large CP of the Independent Subpopulations model reflects the extreme confounding and estimation bias for this data, for which the PC-Lasso and LMM-Lasso methods are able to correct. Meanwhile, the less dramatic improvement in MSE for the 1D Linear Admixture data by the PC-Lasso and LMM-Lasso methods confirms what we know from this data's substantially smaller CP value; that there is less confounding and bias which can be corrected for.

Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical CP with those from the simulated data sets is more complex... this increases with equal subpopulation sizes

\anna{\textbf{check this value for sampsizes of 588 for admixture and independent...maybe even diff subpop sizes? at least to pin down the issue - more comparable with same sampesize and subpop sizes but still different...}}

 \anna{correlation between columns of X..., mspe vs mse- for various reasons CP is a reasonable measure but not perfect...}

\anna{with the same sample size and subpopulation structure as the vanderwas data, I can replicate the CP for the indepenent subpopulations data when simulating using a lower Fst value (as defined in the ochoa and storey papers: the mean inbreeding coefficient of the individuals’ local populations relative to their last common ancestral population. This FST naturally yields a useful pairwise FST between individuals.) So, this suggests that the greater improvement in the semi-empirical data is because there is less inbreeding (?????) still thinking about this...}

@@

\begin{table}[H]
\centering
\begin{tabularx}{400pt}{cclYYY}
\toprule
$\eta$ & $\xi$ & \multicolumn{1}{c}{Method} & 1D Linear Admixture & Independent Subpopulations & Empirical \\ 
\midrule
\multirow{6}{*}{0.5} & \multirow{3}{*}{0} & Lasso & 0.25 (0.05) & 0.23 (0.05) & 0.37 (0.06) \\ 
& & PC-Lasso & 0.23 (0.05) & 0.21 (0.05) & 0.36 (0.06) \\ 
& & LMM-Lasso & 0.25 (0.05) & 0.23 (0.05) & 0.41 (0.06) \\
\cmidrule{2-6}
& \multirow{3}{*}{0.8} & Lasso & 0.23 (0.05) & 0.25 (0.05) & 0.29 (0.06) \\ 
& & PC-Lasso & 0.28 (0.05) & 0.36 (0.06) & 0.47 (0.06) \\ 
& & LMM-Lasso & 0.29 (0.05) & 0.39 (0.06) & 0.58 (0.06) \\ 
\bottomrule
\end{tabularx}
\caption{True positive rates, mean (SD), for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
\label{tab:var_sel}
\end{table}

@@

Given the superior performance of PC-Lasso and LMM-Lasso compared to the ordinary Lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures. Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$. Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations simulated $\bX$ with environmental confounding generated using a dichotomous-discordant $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\bgamma$ structures. The y-axis displays the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures. The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy.

In the presence of coarse population structure, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding boxes straddling the 0 line. However, the majority of these boxes fall above 0, particularly for lower $\eta$ values, indicating the superior performance of LMM-Lasso. In the presence of fine population structure, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy. This is consistent with \citet{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models. As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform reasonably well in the coarse subpopulation setting, but fall short in the presence of fine structure. 
\begin{figure}[H]
    \centering
    % \includegraphics[scale = 0.9]{figures/mse_diff_subpops}
     \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations data, $\eta = 0.5$, and dichotomous-discordant environmental effects.}
    \label{fig:big_vs_small}
\end{figure}

@@

In addition to evaluating PC-Lasso and LMM-Lasso in the presence of coarse and fine subpopulation structure, we also consider distinct patterns of environmental effects. Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) broken down by both subpopulation and coefficient structure. Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and the Independent Subpopulations simulated $\bX$. The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance. The upper left hand panel of Figure \ref{fig:big_small_gamma} corresponds to coarse subpopulation structure with discordant coefficients and also appears to be a setting where LMM-Lasso outperforms PC-Lasso, regardless of coefficient magnitude, although there is some overlap with PC-Lasso. The lower left hand panel of Figure \ref{fig:big_small_gamma} shows results for the setting of coarse supopulation structure with concordant coefficients. In this setting, PC-Lasso and LMM-Lasso perform comparably in the presence of exponential-concordant environmental effects, while PC-Lasso outperforms LMM-Lasso in the majority of dichotomous-concordant environmental effects. These findings are in line with our initial hypothesis that PC-Lasso will perform well in the presence of environmental effects that can be well-approximated by a linear combination of the observed data. When that is not the case, either due to the presence of fine subpopulation structure, nonlinear environmental effects, or genetic similarity that does not reflect similar environmental effects, LMM-Lasso exhibits estimation accuracy superior to that of PC-Lasso.

@@

We have detailed the assumptions made by PC-Lasso and LMM-Lasso, which model confounding influences as fixed and random effects, respectively. We have also described the ways in which simulation schemes to evaluate these methods may fail to accurately capture the effects of environmental heterogeneity. In turn, we propose and utilize a simulation approach in which the estimation bias introduced by environmental heterogeneity may be evaluated. We then compare the estimation accuracy of an ordinary Lasso, PC-Lasso, and LMM-Lasso in various confounding scenarios.

Unsurprisingly, we find that PC-Lasso and LMM-Lasso outperform the ordinary Lasso whenever environmental confounding effects are present. However, in the absence of such effects, PC-Lasso may hinder estimation accuracy and perform more poorly than even the ordinary Lasso. This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability. Additionally, we find that LMM-Lasso outperforms PC-Lasso in the presence of fine subpopulation structure and when confounding effects are nonlinear and discordant in nature. In the presence of coarse subpopulation structure which corresponds to concordant, dichotomous environmental effects, PC-Lasso may outperform LMM-Lasso in estimation accuracy. However, LMM-Lasso is more robust to subpopulation structure as well as nonlinear and discordant environmental effects. Therefore, LMM-Lasso is a more robust method for practitioners seeking to obtain unbiased estimates when the exact nature of the population structure and corresponding environmental heterogeneity are unknown.

We considered two implementations of the LMM-Lasso method: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, which differ in their variance estimation procedures and computational implementations. Although we found that both methods perform similarly in terms of $\bbeta$ estimation accuracy, LMM-Lasso-ggmix has several drawbacks. Firstly, it suffers from estimation instability in $n \ge p$ settings. Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably when $n << p$, LMM-Lasso-ggmix requires up to 60 times the computing time of LMM-Lasso-Rakitsch in these settings due to its iterative estimation procedure. The additional time required of LMM-Lasso-ggmix may be worthwhile if the objective of an analysis is to accurately estimate the narrow-sense heritability in $n << p$ settings. However, the LMM-Lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and thus considered by the authors to be more broadly applicable.

\citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmagg$ and $\delta := \sigmaee / \sigmagg$ as previously discussed, however, we have implemented the LMM-Lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$ in the \texttt{R} package, \texttt{penalizedLMM}, available on the author's github repository \url{https://github.com/areisett/penalizedLMM}.

As an additional point of interest, recall that we provide qualitative and mathematical explanations of how the degree to which differential environmental exposures may confound the genotype-phenotype relationship is dependent upon its similarity to population structure. While the objective of these explanations, and this work as a whole, is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, it also highlights the manner in which such unobserved environmental effects have the potential to inflate residual model variance and bias narrow-sense heritability estimates. A more detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper. However, we hope this emphasizes the broad implications and importance of the subtle distinction between the effect of population structure and unobserved environmental heterogeneity.

@@

\anna{HERITABILITY PARAGRAPH}\\
Although we focus our attention on accurately estimating fixed SNP effects, we note that the result shown in \ref{eqn:partition_e} could lead to substantial bias in narrow-sense heritability estimates. \pb{Hmm...could this sentence be moved somewhere else?  I feel that narrow-sense heritability deserves a paragraph somewhere}

>>>>

 To evaluate the performance of penalized LMM methods in the presence of these different forms of confounding complexity, we consider environmental effects which are concordant or discordant in nature in the presence of both exponential and dichotomous effect sizes, for a total of four distinct $\bgamma$ patterns. Figure \ref{fig:gamma_structures} provides a visual summary of the four coefficient patterns, where the subpopulations on the x-axis are ordered by genetic similarity so that subpopulation 1 is most genetically similar to subpopulation 2, and least similar to subpopulation 10. This may also be thought of as subpopulations ordered spatially along a 1-dimensional geographic region.
\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.9]{figures/gamma_structure.png}
    \caption{An illustration of four distinct patterns of environmental effects and corresponding $\bgamma$ values in the presence of 10 subpopulations ordered by genetic similarity.}
    \label{fig:gamma_structures}
\end{figure}
\noindent In the case of concordant and dichotomous effects, Figure \ref{fig:gamma_structures} shows that the first five ordered subpopulations are subject to an effect size of 1, while the second five are subject to an effect size of $-$1. In the case of discordant dichotomous effects, all subpopulations again experience an effect size with an absolute value of 1, however, the directions of these effects alternate among the ordered subpopulations so that genetically similar subpopulations do not experience similar environmental effects.

@@

We begin by comparing the ordinary Lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied. Estimation accuracy for Lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) and broken down into bias and variance components. Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$. The top row of Figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), Lasso and LMM-Lasso perform comparably. PC-Lasso results in a slight inflation in MSE compared to the other methods in the simulated data sets due to the fact that 10 additional predictors (PCs) have been added to the model, but none of them actually explain any variability in the outcome.

The bottom row of Figure \ref{fig:mse} shows the MSE when $\xi = 0.8$. In this case, both PC-Lasso and LMM-Lasso clearly outperform the ordinary Lasso. This improvement is more substantial in the Empirical and Independent Subpopulations data than in the 1D Linear Admixture data. Qualitatively, this is due to the fact that the Independent Subpopulations and Empirical data sets are characterized by more discrete, nonoverlapping genetic subpopulations, which more directly correspond to the nonoverlapping environmental effects than do the continuous and blended subpopulations of the 1D Linear Admixture data. In other words, a greater proportion of the environmental effect is in the column space of $\boldsymbol{P}_{\bX}$ for the Empirical and Independent Subpopulations data than for the 1D Linear Admixture data. We can quantify the proportion of the environmental component in the column space of $\boldsymbol{P}_{\bX}$ for each data type using the CP metric defined in \eqref{eqn:ratio} of Section \ref{sec:sim_env_conf}. CP $= 0.47, 0.99, 0.61$ for the 1D Linear Admixture, Independent Subpopulations, and Empirical data, respectively. The large CP of the Independent Subpopulations model reflects the extreme confounding and estimation bias for this data, for which the PC-Lasso and LMM-Lasso methods are able to correct. Meanwhile, the less dramatic improvement in MSE for the 1D Linear Admixture data by the PC-Lasso and LMM-Lasso methods confirms what we know from this data's substantially smaller CP value; that there is less confounding and bias which can be corrected for.

Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical CP with those from the simulated data sets is not straightforward. We note that the CP decreases in the presence of subpopulations of unequal sizes and with more outbred subpopulations as in the Empirical data.

% \anna{MSPE here (?)}

@@

\begin{table}[H]
\centering
\begin{tabularx}{400pt}{cclYYY}
\toprule
$\eta$ & $\xi$ & \multicolumn{1}{c}{Method} & 1D Linear Admixture & Independent Subpopulations & Empirical \\ 
\midrule
\multirow{6}{*}{0.5} & \multirow{3}{*}{0} & Lasso & 0.25 (0.05) & 0.23 (0.05) & 0.37 (0.06) \\ 
& & PC-Lasso & 0.23 (0.05) & 0.21 (0.05) & 0.36 (0.06) \\ 
& & LMM-Lasso & 0.25 (0.05) & 0.23 (0.05) & 0.41 (0.06) \\ 
\cmidrule{2-6}
& \multirow{3}{*}{0.8} & Lasso & 0.24 (0.05) & 0.26 (0.06) & 0.29 (0.06) \\
& & PC-Lasso & 0.28 (0.05) & 0.36 (0.06) & 0.46 (0.06) \\ 
& & LMM-Lasso & 0.28 (0.05) & 0.39 (0.06) & 0.57 (0.05) \\
\bottomrule
\end{tabularx}
\caption{True positive rates, mean (SD), for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
\label{tab:var_sel}
\end{table}

@@

Given the superior performance of PC-Lasso and LMM-Lasso compared to the ordinary Lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures. Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.

Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations simulated $\bG$ with environmental confounding generated using a dichotomous-discordant $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\bgamma$ structures. The y-axis displays the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures. The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy.

In the presence of coarse population structure, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding boxes straddling the 0 line. However, the majority of these boxes fall above 0, indicating the superior performance of LMM-Lasso. In the presence of fine population structure, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy across all levels of environmental confounding. This is consistent with \citet{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models. As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform reasonably well in the coarse subpopulation setting, but fall short in the presence of fine structure. 
\begin{figure}[H]
    \centering
    % \includegraphics[scale = 0.9]{figures/mse_diff_subpops}
     \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations data, $\eta = 0.5$, and dichotomous-discordant environmental effects.}
    \label{fig:big_vs_small}
\end{figure}

@@

In addition to evaluating PC-Lasso and LMM-Lasso in the presence of coarse and fine subpopulation structure, we also consider distinct patterns of environmental effects. Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) broken down by both subpopulation and coefficient structure. Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and the Independent Subpopulations simulated $\bG$.

The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance. The upper left hand panel of Figure \ref{fig:big_small_gamma} corresponds to coarse subpopulation structure with discordant coefficients and also appears to be a setting where LMM-Lasso outperforms PC-Lasso, regardless of coefficient magnitude, although there is some overlap with PC-Lasso. The lower left hand panel of Figure \ref{fig:big_small_gamma} shows results for the setting of coarse supopulation structure with concordant coefficients. In this setting, PC-Lasso and LMM-Lasso perform comparably in the presence of exponential-concordant environmental effects, while LMM-Lasso outperforms PC-Lasso in the majority of dichotomous-concordant environmental effects. These findings align with our initial hypothesis that PC-Lasso will perform well in the presence of simple environmental effect patterns that can be well-approximated by a linear combination of the observed data. When that is not the case, either due to the presence of fine subpopulation structure, dichotomous effects, or genetic similarity that does not reflect similar environmental effects, LMM-Lasso exhibits estimation accuracy superior to that of PC-Lasso.

@@

We have detailed the assumptions made by PC-Lasso and LMM-Lasso, which model confounding influences as fixed and random effects, respectively. We have also described the ways in which simulation schemes to evaluate these methods may fail to accurately capture the effects of environmental heterogeneity. In turn, we propose and utilize a simulation approach in which the estimation bias introduced by environmental heterogeneity may be evaluated. We then compare the estimation and variable selection accuracy of the ordinary Lasso, PC-Lasso, and LMM-Lasso in various confounding scenarios.

Unsurprisingly, we find that PC-Lasso and LMM-Lasso outperform the ordinary Lasso whenever environmental confounding effects are present. However, in the absence of such effects, PC-Lasso may hinder estimation accuracy and perform more poorly than even the ordinary Lasso. This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability. Additionally, we find that LMM-Lasso outperforms PC-Lasso in the presence of fine subpopulation structure and when confounding effects are discordant in nature. In the presence of coarse subpopulation structure and exponential environmental effects, PC-Lasso performs comparably to LMM-Lasso in estimation accuracy. However, LMM-Lasso is more robust to subpopulation structure and complex patterns of environmental effects. Therefore, LMM-Lasso may be a more robust method for practitioners seeking to obtain unbiased estimates when the nature of the population structure and environmental heterogeneity are unknown.

We considered two implementations of the LMM-Lasso method: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, which differ in their variance estimation procedures and computational implementations. Although we found that both methods perform similarly in terms of $\bbeta$ estimation accuracy, LMM-Lasso-ggmix has several drawbacks. Firstly, it suffers from estimation instability in $n \ge p$ settings. Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably when $n << p$, LMM-Lasso-ggmix requires up to 60 times the computing time of LMM-Lasso-Rakitsch in these settings due to its iterative estimation procedure. The additional time required of LMM-Lasso-ggmix may be worthwhile if the objective of an analysis is to accurately estimate the narrow-sense heritability in $n << p$ settings. However, the LMM-Lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and is considered by the authors to be more broadly applicable.

\citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmagg$ and $\delta := \sigmaee / \sigmagg$ as previously discussed, however, we have implemented the LMM-Lasso-Rakitsch method using the equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$ in the \texttt{R} package, \texttt{penalizedLMM}, available on the author's github repository \url{https://github.com/areisett/penalizedLMM}.

As an additional point of interest, recall that we provide qualitative and quantitative explanations of how the degree to which differential environmental exposures may confound the genotype-phenotype relationship is dependent upon its similarity to population structure. While the objective of these explanations, and this work as a whole, is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, the result shown in \ref{eqn:partition_e} also highlights the manner in which such unobserved and unmodeled environmental effects could lead to substantial bias in narrow-sense heritability estimatess. A more detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper. However, we hope this emphasizes the broad implications and importance of the subtle distinction between the effect of population structure and unobserved environmental heterogeneity.

@@
